### 2024-10-22 paper/paper.qmd
Basic steps of the model

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - David Qi
  - Haobo Ren
  - Xinrui Xie
thanks: "Code and data are available at: https://github.com/HaoboRrrr/USA_Election"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
```

# Introduction

# Data {#sec-data}

## Overview
Josh Pasek have discussed how to the use of Aggregation and Predictive Modeling on polling can increase prediction effectiveness[@Pasek_2015]. In this article, we will make use of predictive modeling to create a model that aim to predict the support rate of Kamala Harris in the 2024 United States presidential election using a linear regression model trained form past polling results and Dow Jones Industrial Average.

## Measurement

## Outcome variables

## Predictor variables

# Model

We will use a linear model to predict the general polling result of Harris.

## Model set-up
For each poll, we assumed they are being sampled similarly. \
Let the support rate for Harris at day $t$ be denoted by $R_t$. And we denote the Dow Jones Industrial Average at day $t$ by $D_t$.
Then we will use linear regression model: 
$$
R_t = D_{t-60} +  D_{t-90} + \epsilon_t
$$
Where $\epsilon_t$ is an independent, normally distributed error term.

## Model justification

We expect a positive relationship between the support rate of Harris and Dow Jones Industrial Average 60 and 90 days before. Since people tend to support the party in power when the economy is better, where the condiction of the economy is reflected by $D_t$.

TODO: find paper to support this.
finishing reading https://dspacemainprd01.lib.uwaterloo.ca/server/api/core/bitstreams/4dfe49d7-72b7-44e5-8fdc-8a5cf0b77b24/content

# Results

# Discussion

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

## Weaknesses and next steps

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

## Diagnostics

\newpage

# Appendix 1: Polling methodology

# Appendix 2: Ideal methodology and survey

# References

### 2024-10-31 scripts/00-simulate_data.R
Update Data Simulation, deleted 2 unused scripts

@@ -1,9 +1,8 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  #state and party that won each division.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Simulates a dataset of polling results, which include polls and the support rate of Kamala Harris and Donald Trump. And simulate DJIA index. 
+# Author: David Qi
+# Date: 30 October 2024
+# Contact: david.qi@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse` package must be installed
 # Any other information needed? Make sure you are in the `starter_folder` rproj
@@ -17,26 +16,47 @@ set.seed(853)
 #### Simulate data ####
 
 
-# Political parties
-candidates <- c("Trump", "Harris", "Kennedy", "Stein", "West")
+n_polls <- 50   # Number of polls
+mean_support_harris <- 45 # Average support for Kamala Harris (%)
+mean_support_trump <- 45  # Average support for Donald Trump (%)
+mean_support_other<- 10 # Average support for Other Candidates added together (%)
+sd_TrumpHarris_poll <- 3             # Standard deviation of random error for Trump and Harris
+sd_Other_poll <- 0.5            # Standard deviation of random error for sum of Other Candidates
+dates <- seq(as.Date("2024-09-12"),as.Date("2024-10-31"),by = "day")
 
-# Create a dataset by randomly assigning states and parties to divisions
-analysis_data <- tibble(
-  division = paste("Division", 1:151),  # Add "Division" to make it a character
-  state = sample(
-    states,
-    size = 151,
-    replace = TRUE,
-    prob = c(0.25, 0.25, 0.15, 0.1, 0.1, 0.1, 0.025, 0.025) # Rough state population distribution
-  ),
-  party = sample(
-    parties,
-    size = 151,
-    replace = TRUE,
-    prob = c(0.40, 0.40, 0.05, 0.1, 0.05) # Rough party distribution
-  )
+
+# Simulate polling results
+poll_results <- data.frame(
+  Poll = 1:n_polls,
+  end_dates <- dates,
+  Harris = pmax(0, pmin(100, rnorm(n_polls, mean = mean_support_harris, sd = sd_TrumpHarris_poll))),
[…]

### 2024-10-31 scripts/01-test_simulated_data.R
Updated Test Simulated script, Deleted old simulated data

@@ -1,9 +1,8 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Tests the structure and validity of the simulated Data
+# Author: David Qi 
+# Date: 30 October 2024
+# Contact: david.qi@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
   # - The `tidyverse` package must be installed and loaded
@@ -14,76 +13,86 @@
 #### Workspace setup ####
 library(tidyverse)
 
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+poll_results <- read_csv("data/00-simulated_data/simulated_data_poll.csv")
+stock_results <- read_csv("data/00-simulated_data/simulated_data_stock.csv")
 
 # Test if the data was successfully loaded
-if (exists("analysis_data")) {
+if (exists("poll_results")) {
   message("Test Passed: The dataset was successfully loaded.")
 } else {
   stop("Test Failed: The dataset could not be loaded.")
 }
 
+if (exists("stock_results")) {
+  message("Test Passed: The dataset was successfully loaded.")
+} else {
+  stop("Test Failed: The dataset could not be loaded.")
+}
 
-#### Test data ####
+#### Test Poll data ####
 
-# Check if the dataset has 151 rows
-if (nrow(analysis_data) == 151) {
-  message("Test Passed: The dataset has 151 rows.")
+# Check if the dataset has 50 rows
+if (nrow(poll_results) == 50) {
+  message("Test Passed: The dataset has 50 rows.")
 } else {
   stop("Test Failed: The dataset does not have 151 rows.")
 }
 
-# Check if the dataset has 3 columns
-if (ncol(analysis_data) == 3) {
-  message("Test Passed: The dataset has 3 columns.")
+# Check if the dataset has 8 columns
+if (ncol(poll_results) == 8) {
+  message("Test Passed: The dataset has 8 columns.")
 } else {
-  stop("Test Failed: The dataset does not have 3 columns.")
+  stop("Test Failed: The dataset does not have 8 columns.")
 }
 
-# Check if all values in the 'division' column are unique
-if (n_distinct(analysis_data$division) == nrow(analysis_data)) {
-  message("Test Passed: All values in 'division' are unique.")
[…]

### 2024-11-01 paper/paper.qmd

@@
 
+## Objective
+The survey aims to precisely capture voter intention and mood across important demographic and geographic divisions in order to predict the result of the US presidential election.
+
+## Sampling Approach
+* Stratified Random Sampling: To replicate the national electorate, stratify the sample according to age, gender, race/ethnicity, education level, urban/rural domicile, and area.
+
+* Sample Size: Aim for a starting sample of five thousand people. This results in around 2000 to 2500 replies every wave, assuming a response rate of 40–50%.
+
+## Respondent Recruitment
+* Panel Providers: Partner with an online panel provider to find respondents that match each stratum’s specifications.Specify quotas to ensure that, for instance, you have 18% of respondents from the Midwest or 25% aging between 30 - 40.
+
+* Random Sampling: Randomly select respondent within each strata.
+
+* Incentives: To promote greater response rates and retention, provide a small monetary reward, such as $5 to $10 for each completed survey.
+
+* Multi-Mode: To improve inclusivity and take into consideration differences in internet availability, use both phone and online surveys.
+
+## Data Validation
+* De-duplication: To keep distinct replies between waves, make sure respondents don't participate more than once each wave.
+
+* Quality checks: highlight straight-lining, inconsistent replies, and other low-effort answers.
+
+## Essential Questions
+* Are you currently registered to vote?
+  - Yes
+  - No
+
+* Which state are you currently registered to vote in?
+  (short answer)
+
+* Do you plan to vote in the upcoming Presidential election?
+  - Yes, I plan to vote in person on election day
+  - Yes, I plan to vote in person early
+  - Yes, I plan to vote by mail
+  - No, I do not plan to vote
+
+* If the election were held today, who would you vote for?
+  - Democrat Kamala Harris
+  - Republican Donald Trump
+  - A third party / Independent candidate
+  - Unsure
+
+* How certain are you about your choice
+  - Scale 1 to 5
+
+* Your gender idetification
+  - Female
+  - Male
+  - Other
+
+* Age Range
+  - 18 to 24
+  - 25 to 29
+  - 30 to 39
+  - 40 to 49
+  - 50 to 59
+  - 60 or older
+
+* Which of the following best describes your racial background?
+  - White or Caucasion
+  - Black or African American
+  - Hispanic
+  - Asian or Pacific Islander
+  - Native American
+  - Other
+
+* How would you describe your educational level
+  - High School or less
+  - College Graduate
+  - Post-graduate degree
+  - Unsure
+
+* Employment Status
+  - Full time
+  - Part time
+  - Unemployed
+
+* What's your estimated household income?
+  - Under $30,000
+  - \$30k - \$49,999
+  - \$50k - \$99,999
+  - \$100k - \$199,999
+  - \$200k - \$249,999
+  - \$250k or more
+  - Unsure
+
+* Do you consider yourself to be lesbian, gay, bisexual, transgender, queer or
+questioning?
+  - Yes
+  - No
+  - Unsure
+  - Prefer not to answer
+
+* Would you best describe the neighborhood or area you live in as
+  - Urban
+  - Suburban
+  - Rural
+  - Unsure
+
+* What is your religious background
+  - Religious
+  - Spiritual
+  - Not religious or spiritual
+
+Link to the survey: https://forms.gle/spJjASENRQ6nKK1E6
+
 # References

### 2024-11-01 scripts/01-test_simulated_data.R
Correct typo

@@ -35,7 +35,7 @@ if (exists("stock_results")) {
 if (nrow(poll_results) == 50) {
   message("Test Passed: The dataset has 50 rows.")
 } else {
-  stop("Test Failed: The dataset does not have 151 rows.")
+  stop("Test Failed: The dataset does not have 50 rows.")
 }
 
 # Check if the dataset has 8 columns

### 2024-11-01 scripts/01-test_simulated_data.R
fix typo

@@ -52,7 +52,7 @@ if (all(abs(poll_results$Harris_reg + poll_results$Trump_reg + poll_results$Othe
   stop("Test Failed: support rate adds up to value other than 100")
 }
 
-# Check if We have stock data for revelent dates.
+# Check if We have stock data for relevant dates.
 if (all(poll_results$end_dates %in% stock_results$dates)) {
   message("Test Passed: We have DJIA index for all relevent dates.")
 } else {

### 2024-11-01 scripts/00-simulate_data.R
Rewrite Download, Clean and test_analysis. Fix some typo in pervious parts.

@@ -15,7 +15,6 @@ set.seed(853)
 
 #### Simulate data ####
 
-
 n_polls <- 50   # Number of polls
 mean_support_harris <- 45 # Average support for Kamala Harris (%)
 mean_support_trump <- 45  # Average support for Donald Trump (%)

### 2024-11-01 scripts/01-test_simulated_data.R

@@ -56,11 +56,11 @@ if (all(abs(poll_results$Harris_reg + poll_results$Trump_reg + poll_results$Othe
 if (all(poll_results$end_dates %in% stock_results$dates)) {
   message("Test Passed: We have DJIA index for all relevent dates.")
 } else {
-  stop("Test Failed: We have DJIA index for all relevent dates.")
+  stop("Test Failed: We don't have DJIA index for all relevent dates.")
 }
 
 # Check if there are any negative values
-if (all(poll_results$Harris_reg> 0) & all(poll_results$Trump_reg> 0 & all(poll_results$Other_reg> 0))) {
+if (all(poll_results$Harris_reg>= 0) & all(poll_results$Trump_reg>= 0 & all(poll_results$Other_reg>= 0))) {
   message("Test Passed: Support rates are non-negative")
 } else {
   stop("Test Failed: There exist nagative support rate.")

### 2024-11-01 scripts/02-download_data.R

@@ -1,13 +1,16 @@
 #### Preamble ####
 # Purpose: Downloads and saves the presidential general election data from https://projects.fivethirtyeight.com/polls/president-general/2024/national/
+# - Downloads and saves the Doe Jones data form https://www.kaggle.com/api/v1/datasets/download/joebeachcapital/dow-jones-and-s-and-p500-indices-daily-update
 # Author: Haobo Ren
 # Date: 19 October 2024
 # Contact: haobo.ren@mail.utoronto.ca
-
+# License: MIT
+# Pre-requisites: 
+# - The `curl` package must be installed and loaded
+# Any other information needed? Make sure you are in the `starter_folder` rproj
 
 
 #### Workspace setup ####
-library(tidyverse)
 library(curl)
 
 #### Download data ####
@@ -18,8 +21,4 @@ url_DJI = "https://www.kaggle.com/api/v1/datasets/download/joebeachcapital/dow-j
 curl_download(url_DJI, "data/01-raw_data/archive.zip")
 # This is a zip file, data needs to be manually unzipped.
 
-# date retrieved: 22 Oct 2024
-
-
-
-         
+# date data is retrieved: 01 November 2024

### 2024-11-01 scripts/03-clean_data.R

@@ -1,10 +1,10 @@
 #### Preamble ####
 # Purpose: Cleans the raw data and extract required data.
 # Author: David QI
-# Date: 22 Oct 2024
+# Date: 01 November 2024
 # Contact: david.qi@mail.utoronto.ca 
 # License: MIT
-# Pre-requisites: data in data/01-raw_data/raw_data.csv, data/01-raw_data/DJIA.csv
+# Pre-requisites: data in data/01-raw_data/raw_data.csv, data/01-raw_data/DJIA.csv (DJIA.csv need to be manually unzipped from archive.zip)
 # Any other information needed? No
 
 #### Workspace setup ####
@@ -13,33 +13,35 @@ library(ggplot2)
 library(tidyverse)
 library(janitor)
 library(lubridate)
-library(broom)
-library(modelsummary)
-library(rstanarm)
-library(splines)
-library(dplyr)
-library(tidyr)
 # TODO: check if all these libraries are needed
 
 
 #### Clean data ####
 data <- read_csv("data/01-raw_data/raw_data.csv") |> clean_names()
 
-# Filter data to Harris estimates based on high-quality polls after she declared
+# Filter data to Harris and Trump estimates based on high-quality polls after Harris declared entering
 just_harris_high_quality <- data |>
   filter(
     candidate_name == "Kamala Harris",
-    numeric_grade >= 2.7 # Need to investigate this choice - come back and fix. 
-    # Also need to look at whether the pollster has multiple polls or just one or two - filter out later
+    numeric_grade >= 2.5
   ) |>
   mutate(
-    state = if_else(is.na(state), "National", state), # Hacky fix for national polls - come back and check
+    state = if_else(is.na(state), "National", state), # fix for national polls
     end_date = mdy(end_date)
   ) |>
-  filter(end_date >= as.Date("2024-07-21")) |> # When Harris declared
+  filter(end_date >= as.Date("2024-07-21")) # When Harris declared
+
+
+# filter out the same set of questionnaires for Trump 
+just_trump_high_quality <- data |>
+  filter(
+    candidate_name == "Donald Trump",
+    question_id %in% just_harris_high_quality$question_id
+  ) |>
   mutate(
-    num_harris = round((pct / 100) * sample_size, 0) # Need number not percent for some models
-  )
+    state = if_else(is.na(state), "National", state), # fix for national polls
+    end_date = mdy(end_date)
[…]

### 2024-11-01 scripts/05-model_data.R
Updated Scripts for modeling, choosed one best model

@@ -1,0 +1,104 @@
+#### Preamble ####
+# Purpose: Test various models and choose one to save
+# Author: David Qi
+# Date: 22 Oct 2024
+# Contact: david.qi@mail.utoronto.ca 
+# License: MIT
+# Pre-requisites: data in "data/02-analysis_data/DJIA_filled.csv", "data/02-analysis_data/just_harris_high_quality.csv", "data/02-analysis_data/just_trump_high_quality.csv"
+# Any other information needed? No
+
+
+#### Workspace setup ####
+library(tidyverse)
+
+#### Read data ####
+harris_data <- read_csv("data/02-analysis_data/just_harris_high_quality.csv") %>% select(pct, start_date,end_date)
+trump_data <- read_csv("data/02-analysis_data/just_trump_high_quality.csv") %>% select(pct, start_date,end_date)
+DJIA_data = read_csv("data/02-analysis_data/DJIA_filled.csv")
+
+### Model data ####
+# prepare data
+
+# Convert the date columns to Date format
+harris_data$start_date <- as.Date(harris_data$start_date, format = "%m/%d/%y")
+harris_data$end_date <- as.Date(harris_data$end_date, format = "%m/%d/%y")
+
+# Create a data frame to store merged data
+
+DJIA_data = DJIA_data %>%  # Create lagged variables
+  mutate(DJIA_lag_1 = lag(DJIA, n = 1),
+         DJIA_lag_7 = lag(DJIA, n = 7),
+         DJIA_lag_14 = lag(DJIA, n = 14),
+         DJIA_lag_28 = lag(DJIA, n = 28),
+         DJIA_lag_60 = lag(DJIA, n = 60),
+         DJIA_lag_90 = lag(DJIA, n = 90),
+         DJIA_lag_180 = lag(DJIA, n = 180),
+  )
+
+merged_harris_data <- harris_data %>%
+  # Left join with DJIA_data based on dates, shifting for lag
+  left_join(DJIA_data, by = c("end_date" = "DATE")) 
+
[…]

### 2024-11-01 scripts/03-clean_data.R
Clean unused library, add used library to reference

@@ -9,12 +9,8 @@
 
 #### Workspace setup ####
 
-library(ggplot2)
 library(tidyverse)
 library(janitor)
-library(lubridate)
-# TODO: check if all these libraries are needed
-
 
 #### Clean data ####
 data <- read_csv("data/01-raw_data/raw_data.csv") |> clean_names()

### 2024-11-04 paper/paper.qmd

@@
 ## Objective
-The survey aims to precisely capture voter intention and mood across important demographic and geographic divisions in order to predict the result of the US presidential election.
+The survey aims to capture voter intention and mood across important demographic and geographic divisions in order to predict the result of the US presidential election.
+This survey's methodology employs stratify sampling, effective respondent recruitment strategy and data validation techniques to ensure the accuracy of prediction.
 
 ## Sampling Approach
-* Stratified Random Sampling: To replicate the national electorate, stratify the sample according to age, gender, race/ethnicity, education level, urban/rural domicile, and area.
-
-* Sample Size: Aim for a starting sample of five thousand people. This results in around 2000 to 2500 replies every wave, assuming a response rate of 40–50%.
+This survey employs Stratified Random Sampling, which involves divide the target population(In this case, the people who are eligible to vote in US) into subgroups(strata) based on their demographics characteristic.
+To simulate the national electorate, stratify the sample according to age, gender, race/ethnicity, education level, urban/rural domicile, and region.
+Also refer to the US Census data to determine the proportion of each strata, for example: if the census data shows there are 20% US citizen are between the age of 20 to 30, then there should be 20% of respondent in this age range.
+We aim for a starting sample of five thousand people since there were budget constraint.
 
 ## Respondent Recruitment
-* Panel Providers: Partner with an online panel provider to find respondents that match each stratum’s specifications.Specify quotas to ensure that, for instance, you have 18% of respondents from the Midwest or 25% aging between 30 - 40.
-
-* Random Sampling: Randomly select respondent within each strata.
-
-* Incentives: To promote greater response rates and retention, provide a small monetary reward, such as $5 to $10 for each completed survey.
-
-* Multi-Mode: To improve inclusivity and take into consideration differences in internet availability, use both phone and online surveys.
+Partner with an online panel provider to find respondents that match each stratum’s specifications.Specify quotas to ensure that, for instance, you wish to have 18% of respondents from the Midwest or 25% aging between 30 - 40.
+Then randomly select respondent within each strata.
+To promote greater response rates and retention, provide a small monetary reward, such as $5 for each completed survey.
+To improve inclusivity and take into consideration differences in internet availability, use both phone and online surveys.
 
@@
 
-## Essential Questions
+## Budget Expenditure
+* \$50k - \$60k for panel provider
+* \$25k for incentives
+* \$10k for Telephone survey services
+* \$5k for data analysis
+
+## Survey Structure
+This survey will focus only on collecting participants' preferences for presidential candidates, not senate elections.Therefore, this survey only asks participants' choice of their favored presidential candidates.
+This design can make the entire questionnaire more simplify, while also collecting key data for predicting the presidential election.
+
+Screener Section: Screening out people who are eligible to vote in each state. This can filter out participants who are not eligible to vote, thereby improving the accuracy of the prediction results.
+
 * Are you currently registered to vote?
@@
 * Which state are you currently registered to vote in?
-  (short answer)
+  - (Drop down box)
+
+Presidential Ballot Section: Ask participants whether they would vote, who they would vote for, and how confident they are about that.
+This section can be used to find out the proportion of votes each candidate received among the participants.
 
@@
 
-* Your gender idetification
+Demographics Section: This section collects detailed demographic characteristics of the participants to accurately predict the nation-wide election outcome.
+
+* Your gender identification
   - Female
@@
 
-* Which of the following best describes your racial background?
-  - White or Caucasion
+* Which of the following best describes your race or ethnicity
+  - White or Caucasian
   - Black or African American

### 2024-11-04 paper/paper.qmd
Add more details to the Appendix 2: Ideal methodology and survey, and more citations to the references.bib

@@
 ## Sampling Approach
-This survey employs Stratified Random Sampling, which involves divide the target population(In this case, the people who are eligible to vote in US) into subgroups(strata) based on their demographics characteristic.
+This survey employs Stratified Random Sampling, which involves divide the target population(In this case, the people who are eligible to vote in US) into subgroups(strata) based on their demographics characteristic[@stantchevaonline].
 To simulate the national electorate, stratify the sample according to age, gender, race/ethnicity, education level, urban/rural domicile, and region.
@@
 ## Respondent Recruitment
-Partner with an online panel provider to find respondents that match each stratum’s specifications.Specify quotas to ensure that, for instance, you wish to have 18% of respondents from the Midwest or 25% aging between 30 - 40.
+Partner with an online panel provider to find respondents that match each stratum’s specifications[@stantchevaonline].
+Specify quotas to ensure that, for instance, you wish to have 18% of respondents from the Midwest or 25% aging between 30 - 40.
 Then randomly select respondent within each strata.
@@
 ## Data Validation
-* De-duplication: To keep distinct replies between waves, make sure respondents don't participate more than once each wave.
-
-* Quality checks: highlight straight-lining, inconsistent replies, and other low-effort answers.
+To keep distinct replies between responses, make sure respondents don't participate more than once each wave, and highlight straight-lining, inconsistent replies, and other low-effort answers for data quality purposes[@horn1997effective]
 
@@
 ## Survey Structure
-This survey will focus only on collecting participants' preferences for presidential candidates, not senate elections.Therefore, this survey only asks participants' choice of their favored presidential candidates.
-This design can make the entire questionnaire more simplify, while also collecting key data for predicting the presidential election.
 
-Screener Section: Screening out people who are eligible to vote in each state. This can filter out participants who are not eligible to vote, thereby improving the accuracy of the prediction results.
+### Introduction:
+
+First of all, thank you for taking the time to participate in this survey.
+
+This survey aims to precisely capture voter intention and mood across important demographic and geographic divisions in order to predict the result of the US presidential election.
+
+Please note:
+
+  - Your responses will remain confidential.
+
+  - Please answer the question honestly.
+
+  - Complete the survey will receive a reward for \$5 - \$10.
+
+If you have concerns or questions, reach out to haobo.ren@mail.utoronto.ca (Haobo Ren)
+
+### Screener Section: Screening out people who are eligible to vote in each state.
 
@@
 
-Presidential Ballot Section: Ask participants whether they would vote, who they would vote for, and how confident they are about that.
-This section can be used to find out the proportion of votes each candidate received among the participants.
+### Presidential Ballot Section: Ask participants whether they would vote, who they would vote for, and how confident they are about that.
 
@@
 
-Demographics Section: This section collects detailed demographic characteristics of the participants to accurately predict the nation-wide election outcome.
+### Demographics Section: This section collects detailed demographic characteristics of the participants.
 

### 2024-11-04 paper/paper.qmd
Edit Appendix 1: Polling methodology With details of the reference list and in text citations

@@
 
+## Overview of Morning Consult
+Morning Consult is a globally renowned enterprise technology company that specializes in providing intelligent data to support leaders' decision-making processes.
+Morning Consult has established itself as a key player in the field of artificial intelligence and decision-making consulting.
+The company offers a suite of services that include market research, brand analysis, consumer behavior studies, and more.
+So that the company can optimize decision-making processes and provide clients with a competitive edge[@morningconsult_website]
+
+## Features of the Sample
+
+  - Population: The survey aims at all the American citizens who is eligible to vote in the 2024 presidential election in all the parties.
+
+  - Sampling Frame: A list of the American citizen’s email address and telephone number.
+
+  - Sample Size: The polling samples showed a time range from 2011 to 2024.
+During each election period, the sample size varies a lot.
+The sample size ranges from 111 to 78247 across the voting period.
+
+## Sample Recruitment
+
+The company used many sampling methods, such as online panel, live phone call, text, emails and a mix of these approaches.
+During the recruitment, the data was collected by contacting landlines via Interactive Voice Response.
+The sampling teams made phone calls and sent text message to randomly selected phone numbers, and collect their responses.
+The online team collect votes from the active users in the websites and other online panel of voters provided by research marketplace such as CINT.
+However, some of the polling results come from a survey from long ago, which means that it might be incorrect or biased[@morningconsult_poll].
+
+## Sample Approach
+
+Both probability and non-probability sampling methods are involved in the sample collection.
+For the people surveyed via phone calls and text messages, the main sampling method is Simple Random Sample (SRS).
+While for those sampled via online panel, the main approach is convenience sample.
+
+The Morning Consult company used so-called "tracking polls" [@morningconsult_poll], interviews from a certain period are incorporated into future polls, getting reweighted with different samples until they are too old and dropped from the analysis.
+
+## Strengths and Weakness
+
+Strengths: Since both probability and non-probability sampling methods are used, the pollster can cover a wider range of people surveyed.
+This means that the sample result can better represent the whole population.
+Also, the reweighting mechanism can guarantee different methods can have different influence on the final result.
+
+Weakness: According to the data collected, majority of the pollster are done by the convenience sample, which is a non-probability sampling.
+Thus the actual surveyed population may be biased from the target population, and the result may be less representative.
+Additionally, the survey showed some overlapping time period, which means that some opinions of the interviewers may be counted more than once.
+
+## Non-response Handling
+
+Just like all the surveys, non-response bias is the biggest challenge met during the pollster.
+This is also the greatest factor that may affect the effectiveness of the survey.
+In all these sampling recruitment and sampling approaches, a portion of the respondents would fail to answer the survey questions as required for various reasons (such as refusing to participate, being unreachable, or being unable to understand the survey questions).
+
+Based on the online research, Morning Consult company usually adopt these measures to minimize the effect of the non-response bias, such as increasing the survey response rate, following up with non-respondents, and using a broader sample for the survey.
+Additionally, by comparing and analyzing the characteristics of respondents and non-respondents and supplementing the data with other relevant sources, researchers can further assess the influence of non-response bias on the study results.
+
+According to the collected data, we can also see that all the sample recruitment methods request the use of the cell phone.
+This automatically left out the people who do not use cell phones or who do not use Internet, such as some old people.
+But this type of non-response bias is still unsolved.
+
+## Comments of the Questionnaire
+
+The good thing about the questionnaires is that most of the questions are direct and easy to understand.
+So that this can avoid the bias due to misunderstanding of the questions.
+Secondly, the questionnaire is short, which makes more people willing to answer without causing waste of time.
+
+On the other hand, the questionnaire requests the interviewers to identify their party, which may cause people not willing to vote for the candidates from the other party, even though they actually want to.
+This will also cause a slight bias in the response.
+
 # Appendix 2: Ideal methodology and survey
@@
 ## Data Validation
-To keep distinct replies between responses, make sure respondents don't participate more than once each wave, and highlight straight-lining, inconsistent replies, and other low-effort answers for data quality purposes[@horn1997effective]
+To keep distinct replies between responses, make sure respondents don't participate more than once each wave, and highlight straight-lining, inconsistent replies, and other low-effort answers for data quality purposes[@horn1997effective].
 

### 2024-11-04 scripts/02-download_data.R
Modified scripts: Use parquet, modified model for better interpretation

@@ -20,5 +20,6 @@ download.file(url, destfile = "data/01-raw_data/raw_data.csv")
 url_DJI = "https://www.kaggle.com/api/v1/datasets/download/joebeachcapital/dow-jones-and-s-and-p500-indices-daily-update"
 curl_download(url_DJI, "data/01-raw_data/archive.zip")
 # This is a zip file, data needs to be manually unzipped.
+# From kaggle, however the original data was retrived from FRED(https://fred.stlouisfed.org/series/DJIA)
 
 # date data is retrieved: 01 November 2024

### 2024-11-04 scripts/03-clean_data.R

@@ -11,21 +11,54 @@
 
 library(tidyverse)
 library(janitor)
+library(arrow)
 
 #### Clean data ####
 data <- read_csv("data/01-raw_data/raw_data.csv") |> clean_names()
+DJIA_data <- read_csv("data/01-raw_data/DJIA.csv")
+
+### Generate State summary  $$$$
+just_harris_data = data%>% filter(
+     candidate_name == "Kamala Harris" )  %>% mutate(
+         state = if_else(is.na(state), "National", state), # fix for national polls
+         end_date = mdy(end_date)
+       ) |>
+     filter(end_date >= as.Date("2024-07-21"))
+
+harris_states <-  just_harris_data %>% group_by(state) %>% 
+     summarise(total_count=n(),
+                             mean_pct = mean(pct),
+                             .groups = 'drop') 
+ 
+just_trump_data <- data%>% filter(
+       candidate_name == "Donald Trump", )  %>% mutate(
+           state = if_else(is.na(state), "National", state), # fix for national polls
+           end_date = mdy(end_date)
+         ) |>
+    filter(end_date >= as.Date("2024-07-21"))
+
+trump_states = just_trump_data %>% group_by(state) %>% 
+     summarise(total_count=n(),
+              mean_pct = mean(pct),
+              .groups = 'drop')
+
 
 # Filter data to Harris and Trump estimates based on high-quality polls after Harris declared entering
 just_harris_high_quality <- data |>
   filter(
     candidate_name == "Kamala Harris",
-    numeric_grade >= 2.5
+    numeric_grade >= 2.5,
   ) |>
   mutate(
     state = if_else(is.na(state), "National", state), # fix for national polls
     end_date = mdy(end_date)
   ) |>
-  filter(end_date >= as.Date("2024-07-21")) # When Harris declared
+  filter(end_date >= as.Date("2024-07-21"), # When Harris declared 
+         
+         state %in% c("North Carolina", "Pennsylvania", "Georgia", "Nevada",
+                      "Arizona", "Michigan", "Wisconsin", 
+                      "National", "Maine CD-2", "Florida") # relevant states
+         ) 
 
 
 # filter out the same set of questionnaires for Trump 
[…]

### 2024-11-04 scripts/04-test_analysis_data.R

@@ -4,16 +4,17 @@
 # Date: 01 November 2024
 # Contact: david.qi@mail.utoronto.ca 
 # License: MIT
-# Pre-requisites: DJIA_filled.csv, just_trump_high_quality.csv, just_harris_high_quality.csv are in data/02-analysis_data
+# Pre-requisites: DJIA_filled.parquet, just_trump_high_quality.parquet, just_harris_high_quality.parquet are in data/02-analysis_data
 # Any other information needed? No.
 
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
 
-harris_data <- read_csv("data/02-analysis_data/just_harris_high_quality.csv") %>% select(pct, start_date,end_date,question_id)
-trump_data <- read_csv("data/02-analysis_data/just_trump_high_quality.csv") %>% select(pct, start_date,end_date,question_id)
-DJIA_data = read_csv("data/02-analysis_data/DJIA_filled.csv")
+harris_data <- read_parquet("data/02-analysis_data/just_harris_high_quality.parquet") %>% select(pct, start_date,end_date,question_id,state)
+trump_data <- read_parquet("data/02-analysis_data/just_trump_high_quality.parquet") %>% select(pct, start_date,end_date,question_id,state)
+DJIA_data = read_parquet("data/02-analysis_data/DJIA_filled.parquet")
 
 # Test if the data was successfully loaded
 if (exists("harris_data")) {
@@ -57,6 +58,15 @@ if (!any(duplicated(harris_data))) {
   stop("Test Failed: The Harris dataset contains duplicates.")
 }
 
+# Check if data only include relevant states
+if (all(harris_data$state %in% c("North Carolina", "Pennsylvania", "Georgia", "Nevada",
+                                "Arizona", "Michigan", "Wisconsin", 
+                                "National", "Maine CD-2", "Florida"))){
+  message("Test Passed: Harris Data set only contain data for relevant states")
+}else{
+  stop("Test Failed: Harris data set have for states that we are not modeling")
+}
+
 #### Test Trump data ####
 
 # Check if there are any negative values
@@ -80,6 +90,15 @@ if (!any(duplicated(trump_data))) {
   stop("Test Failed: The Trump dataset contains duplicates.")
 }
 
+# Check if data only include relevant states
+if (all(trump_data$state %in% c("North Carolina", "Pennsylvania", "Georgia","Nevada", 
+                                "Arizona", "Michigan", "Wisconsin", 
+                                "National", "Maine CD-2", "Florida"))){
[…]

### 2024-11-04 scripts/05-model_data.R

@@ -10,11 +10,12 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
 
 #### Read data ####
-harris_data <- read_csv("data/02-analysis_data/just_harris_high_quality.csv") %>% select(pct, start_date,end_date)
-trump_data <- read_csv("data/02-analysis_data/just_trump_high_quality.csv") %>% select(pct, start_date,end_date)
-DJIA_data = read_csv("data/02-analysis_data/DJIA_filled.csv")
+harris_data <- read_parquet("data/02-analysis_data/just_harris_high_quality.parquet") %>% select(pct, start_date,end_date,question_id,state)
+trump_data <- read_parquet("data/02-analysis_data/just_trump_high_quality.parquet") %>% select(pct, start_date,end_date,question_id,state)
+DJIA_data = read_parquet("data/02-analysis_data/DJIA_filled.parquet")
 
 ### Model data ####
 # prepare data
@@ -32,6 +33,8 @@ DJIA_data = DJIA_data %>%  # Create lagged variables
          DJIA_lag_28 = lag(DJIA, n = 28),
          DJIA_lag_60 = lag(DJIA, n = 60),
          DJIA_lag_90 = lag(DJIA, n = 90),
+         DJIA_lag_120 = lag(DJIA, n = 120),
+         DJIA_lag_150 = lag(DJIA, n = 150),
          DJIA_lag_180 = lag(DJIA, n = 180),
   )
 
@@ -42,24 +45,27 @@ merged_harris_data <- harris_data %>%
 # Remove rows with NA values (may occur due to lagging)
 merged_harris_data <- na.omit(merged_harris_data)
 
+# Model
+
 # Fit the linear regression model
[…]

### 2024-11-04 paper/paper.qmd
Edit README.md and change of the paper's formating
Update paper: Main body section

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Analysing the 2024 U.S. Presidential Election"
+subtitle: "Spotlight on North Carolina's Role as a Battleground State"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper examines the relationship between polling results and economic conditions to forecast the 2024 U.S.
+presidential election, with results analyzed for both swing states and at a national level.
+Findings suggest that Republicans have a modest chance (6.6%) of winning, with North Carolina emerging as the pivotal swing state.
+The study may offer guidance for campaign strategies and market forcasts.
+While providing insight on changes of swing states in U.S.
+politics.
+"
 format: pdf
 number-sections: true
+toc: true
 bibliography: references.bib
@@
 library(tidyverse)
+library(arrow)
+library(knitr)
+library(ggplot2)
+
+harris_data <- read_parquet(here::here("data/02-analysis_data/just_harris_high_quality.parquet")) %>% select(pct, start_date,end_date,question_id,state)
+trump_data <- read_parquet(here::here("data/02-analysis_data/just_trump_high_quality.parquet")) %>% select(pct, start_date,end_date,question_id,state)
+DJIA_data = read_parquet(here::here("data/02-analysis_data/DJIA_filled.parquet"))
+
+merged_summary = read_parquet(here::here("data/02-analysis_data/merged_summary.parquet"))
+
+model_harris = readRDS(file = here::here("models/model_harris.rds"))
+
+model_trump = readRDS(file = here::here("models/model_trump.rds"))
+
+prediction_results = read_parquet(here::here("models/prediction_results.parquet"))
+
+trump_data <- read_parquet(here::here("data/02-analysis_data/just_trump_high_quality.parquet"))
+
 ```
 
-# Introduction
+# Introduction {#sec-intro}
+
+Josh Pasek have discussed how to the use of Aggregation and Predictive Modeling on polling can increase prediction effectiveness[@Pasek_2015].
+Following this principle, we will make use of predictive modeling to create a model that aim to predict the results of the 2024 United States presidential election.
+This paper uses polling data gathered by FiveThirtyEight[@fivethirtyeight] with Dow Jones industrial average data[@sp_dow_jones_2024] as an indicator of the economy to train a simple linear regression model to predict the result of election in each swing state.
+
+The model aims to predict the result of a poll that ends on November $5^{th}$, 2024, we believe this will reflect the result of the upcoming election on that day, the result of the election may be assumed to be close to the mean of these polls.
+And by this process we derive the result of the election.
+After the result is gained, we computed the electoral votes for the state, and used Monte Carlo simulation to derive winning possibility of each party.
+
+We conclude that Donald Trump have $6.6\%$ probability of winning this election.
+From our prediction, we highlight North Carolina as the most important battleground state.
+The state has a predicted $14.2\% $probability to be won by Trump, and will give him 16 electoral votes.
+Along with the votes that we predict he will almost certainly gain, this pushes him one step (9 electoral votes) to victory.
+This provide power insight to how campaign and fund raising strategies should be targeted on the election day.
+And provides a forecast on one of the most important influences in economy in the next several months, which will be especially helpful in planning market stragities.
+
+The scripts in this paper are written in R[@citeR], The table in our study are generated by knitr [@knitr],  and graph with ggplot2[@ggplot2].
+Data input and output were done using arrow[@arrow], and curl[@curl].
+Data cleaning made use of janitor[@janitor].
+
+The rest of the paper is structured as follows: in [Data](#sec-data) section we will discuss the various aspects of the data we use, how they were gathered and the variable we shall use.
+In the [Model](#sec-model) section we will discuss details of our model, why it is justified and the coefficient summaries.
+In the [Results](#sec-results) section, we present the predictions of the model.
+In the [Discussion](#sec-discussion) section we will discuss the interpretation to the predictions, and what we may learn from the coefficients for the model, we will also look as possible weakness and future improvement for the model.
 
 # Data {#sec-data}
+We will make use of polling data organized by FiveThirtyEight[@fivethirtyeight], collected from polling made by various sources before the election.
+As well as the Dow Jones industrial average data[@sp_dow_jones_2024].
 
-## Overview
-Josh Pasek have discussed how to the use of Aggregation and Predictive Modeling on polling can increase prediction effectiveness[@Pasek_2015].
-In this article, we will make use of predictive modeling to create a model that aim to predict the support rate of Kamala Harris in the 2024 United States presidential election using a linear regression model trained form past polling results and Dow Jones Industrial Average.
+## Overview For Polling data
+Fist we take a look at the the polling data.
 
-## Measurement
+The data set contain 17208 observations of polls, for our propose, we choose only to include 1176 entries, the justification for this data selection is given in the [Justification](#justification_poll) section.
 
-## Outcome variables
+##Data Measurement
+The data was organized by FiveThirtyEight (2024), collected form polls made by various institutions, they each use a distinct methodology.
+While FiveThirtyEight checks their methodology, and include the polling data if it meets their standard.
+An example of the methodology of one poll is included in [Appendix 1](#appendix1) It is a reliable and comprehensive source of US election polls， so we have choose this data set.
 
-## Predictor variables
+## Justification For Poll Selection{#justification_poll}
+The goal of our study is to predict the result of US presidential election, it is known that the only reasonable winner will be form the Democratic Party, or the Republican Party.
+Which can be seen form their overwhelmingly high support rate in the polling results and historical results.
+So we will only include the polling results for Kamala Harris and Donald Trump, the candidate for the two parties.
 
-# Model
+Another observation coming form past experience that is not so obvious is that the result of presidential elections are often only determined by the result in "swing states".[@why_only_ten; @aldrich2023change] Following the definition given by Schultz and Hecht, which have given having below 5% difference between the two parties as one criteria, we have examined our data and have identified the swing states from the data set, past experience, and outside sources.
 
-We will use a linear model to predict the general polling result of Harris.
+[@tbl-polloverview] Shows the absolute value of the difference in average support rate from all the polls.
+If we take all states such the absolute difference is less than 5% to be swing states, similar to the method of less than 5% difference between two parties in the last election.
+We will see they coincide except that our list includes Maine CD-2 and have not included Florida[@fec2020federalelections].
+This list, when Florida is added, also include the swing state listed by popular media[@bbc2024election;@dorn2024election;@abcnews2024trumpleads] and polling platform[@fivethirtyeight2024swing] before this upcoming election.
+
+Also note that some states only have very limited polls, which will make modeling these states very difficult and inaccurate, the swing states we choose allow us to have sufficient polls for each state.
+
+The list of swing states we use are: North Carolina, Pennsylvania, Nevada, Georgia, Arizona, Michigan, Wisconsin, Maine's 2nd congressional district, and Florida.
+We will also include national polls to help with modeling.
+
+The data have also contained ratings for each poll based on their methodology and transparency.
+As a summary made to understand the data, [@tbl-polloverview] is made by selecting all the polling, however, for the rest of this study, we will only use polls that are rated 2.5 or above by FiveThirtyEight to ensure quality of data.
+
+```{r}
+#| label: tbl-polloverview
+#| tbl-cap: Comparison of average polling result for two candiates
+#| include: true
+#| echo: false
+#| warning: false
+#| message: false
+
+merged_summary = merged_summary %>% select(state,total_count.x,mean_pct_harris,mean_pct_trump,abs_diff)  %>% arrange(abs_diff)
+
+kable(merged_summary, col.names = c("State", "Number of Polls" , "Average Support Rate(Harris)","Average Support Rate(Trump)","Absolute difference"))
+```
+
+## Varibles
+This section explains the important variables relevant to our analysis: Percent, End Date, state, numeric grade, Candidate Name and sample size.
+
+### Percent
+Percent (`pct`) is the percentage support rate learned by this poll. This is the main estimated we are targeted to model, the justification of using this as estimand is given in model section.
+
+### End Date
+End Date(`end_date`) is a time variable that specify the date that the poll ends.
+
+### State
+State (`state`) is a categorical variable representing each U.S.
+state or Congressional district that is considered a separate entity in the U.S.
+presidential election.
+Each poll reflect the result of a state, if the state is not specified, the poll is national.
+
+### Numeric grade
+Numeric grade (`numeric_grade`) variable is a measurement of the quality of the poll given by FiveThirtyEight.It is a numeric score from 3 to 0.
+In this study, we will only use polls with Numeric grade higher than 2.5.
+
+### Candidate Name 
+Candidate Name (`candidate_name`) is a categorical variable that specify the name of candidate whose the support rate this poll measure.
+
+### Sample Size
+[@fig-samplesize] is a plot showing the sample size(`sample_size`) of the data we use for modeling.
+The sample size of each poll is 432 minimum and 78247 maximum, the mean sample size is 1497.
+Observe that the sample size is extremely small comparing to the entire population of a state, this observation is important in how we predict using our model.
+
+```{r}
+#| label: fig-samplesize
+#| tbl-cap: Comparison of average polling result for two candiates
+#| include: true
+#| echo: false
+#| warning: false
+#| message: false
+
+trump_data = trump_data %>% filter(sample_size < 10000)
+
+ggplot(trump_data, aes(x = sample_size)) +
+  geom_histogram( binwidth = 200, fill = "steelblue", color = "white", alpha = 0.7,xlim = 10000)+
+  labs(x = "Sample size",
+       y = "Frequency") +
+  theme_minimal()
+```
+
+## Overview Dow Jones Data
+The Dow Jones Industrial Average(DJIA) is a renowned economic indicator that launched in 1896[@spglobal_djia].
+It is used long in economical analysis and is considered "one of the analyst's oldest friend and one of his most useful tool"[@milne1966dow].
+The index is still being used in economic studies today and is considered an important indicator of the economic of the United States[@arendas2018halloween].
+
+The data includes average DJIA data for each weekday, there are no data for weekends and certain holidays that DJIA do not operate.
+For the propose of this analysis, we have filled the data by assigning these dates with DJIA values on the closest date before.
+The data after filling is a numerical value corresponding to each day.
+
+## DJIA Data Measurement
+The data publisher(S&P Dow Jones Indices LLC) have full transparency and reproducible of their methodology[@milne1966dow], DJIA is a weighted average of stock price of 30 firm that are considered important to all aspect of economy except transportation and utilities.
+The firm and corresponding weights are fully transparent.
+As these firms play a key role in the U.S.
+economy and the presidential election, Dow Jones Industrial Average will be a useful predictor for electoral results.
+
+It is also a daily updated, easy to access data that is not subject to change.
+Some other important indicators of economy, such as United States Nonfarm Payrolls, only updates monthly and previous data may be corrected in the future(@bls_employment_situation).
+
+# Model {#sec-model}
+
+We will use simple linear regression model to predict the general polling result of Harris and Trump. The mode will be implemented using R[@citeR].
 
@@
 For each poll, we assumed they are being sampled similarly. \
-Let the support rate for Harris at day $t$ be denoted by $R_t$. And we denote the Dow Jones Industrial Average at day $t$ by $D_t$.
+Let the support rate for Harris in a single poll that end at day $t$ be denoted by $H_t$.
+And we denote the Dow Jones Industrial Average at day $t$ by $D_t$.Denote State by $s$, and the corresponding regression coefficient by $\beta_s$
 Then we will use linear regression model: 
 $$
-R_t = D_{t-60} +  D_{t-90} + \epsilon_t
+H_t = \beta_0 + \beta_1D_{t-7} +  \beta_2D_{t-14}+\beta_3D_{t-28}+\beta_4D_{t-60}+\beta_5D_{t-150} +\beta_6D_{t-180}+ \beta_s + \epsilon_t
 $$
-Where $\epsilon_t$ is an independent, normally distributed error term.
+Where $\epsilon_t$ is an independent, normally distributed error term.  \
+$\beta_0,\beta_1,\beta_2,\beta_3,\beta_4,\beta_5,\beta_6$ are regression coefficients. \
+$\beta_s$ is the corresponding regression coefficient to state $s$. \
+We will predict the election support rate of Harris to be $mean(H_t)$ \
+
+For Donald Trump, we use a similar model, let the support rate in a single poll for Trump at day $t$ be denoted by $T_t$. The the linear regression model can be written as: 
+$$
+T_t = \alpha_0 + \alpha_1D_{t-7} +  \alpha_2D_{t-14}+\alpha_3D_{t-28}+\alpha_4D_{t-60}+\alpha_5D_{t-150} +\alpha_6D_{t-180}+ \alpha_s + e_t
+$$
+Where $e_t$ is an independent, normally distributed error term.  \
+$\alpha_0,\alpha_1,\alpha_2,\alpha_3,\alpha_4,\alpha_5,\alpha_6$ are regression coefficients. \
+$\alpha_s$ is the corresponding regression coefficient to state $s$. \
+We will predict the election support rate of Donald Trump to be $mean(T_t)$ \
 
 ## Model justification
+In deciding this model, we have made many decisions. In this section we justify the importat choices we made during modeling.
 
-We expect a positive relationship between the support rate of Harris and Dow Jones Industrial Average 60 and 90 days before.
-Since people tend to support the party in power when the economy is better, where the condiction of the economy is reflected by $D_t$.
+### Use of DJIA
+The use of economic data to predict result in presidential election can be supported by various studies that shows the result and voting pattern in presidential elections are highly correlated to the economy[@aldrich2023change].
+And DJIA is an important and wide used indicator of the economy.
+We hypothesize that DJIA will have a linear relationship with the support rate.
+We expect the effect economy has on polling results to have a delayed effect, which is the reason for including lagged terms in linear regression.
 
-TODO: find paper to support this.
-finishing reading https://dspacemainprd01.lib.uwaterloo.ca/server/api/core/bitstreams/4dfe49d7-72b7-44e5-8fdc-8a5cf0b77b24/content
+An alternative approach is to use changes in the DJIA to reflect economic fluctuations.
+However, we choose to use the absolute DJIA values, as inputing missing days for change data is difficult to justify, and the absolute economic status provides essential insights.
+Our hypothesis is that a strong economy favors the ruling party in elections, as switching parties could risk economic shifts and potentially lower living standards.
+Here, the absolute economic level serves as an indicator, reflecting the overall standard of living.
 
-# Results
+### Use of mean to predict election {#usemean}
+We may treat each poll as an observation for the population of support rate, what our model does is that it predicts the Percent('pct') variable in a single poll.
+Since the sample size of is extremely small comparing to the population of each state, which is 1497 people on average comparing to at least 700 thousand people.
+And by statistic around 60% people will participate in each election[@aldrich2023change], we may consider the election result to be very close to the population mean by the weak law of large numbers.
+So by considering the mean of the predicted polls, we may predict the election result.
 
-# Discussion
+### Using different model for Harris and Trump
+There exist the possibility for using one model with a categorical variable, however we find the voting behavior of the two groups of people may be vastly different and cannot be summarized by a simple shift in the regression line.
+The use of interaction terms may capture this difference, however will make the model more complicated and difficult to write out.
 
-## First discussion point {#sec-first-point}
+### Using the same model for each state
+As states in United States, we expect each state to share some similarity with other states, as well as with the national data.
+So we have made state a categorical variable in regression.
+The main difference with candidates here is that we have data on national polls, which have direct correlation with the polling result of each state rather than just hypothesized connection, to make use of those data, we include them by using a categorical for state.
 
-## Second discussion point
+### Linerity and Independence Assumption
+The justification for these assumptions are in [Appendix 3](#appendix3)
 
-## Third discussion point
+### Reason for choosing these lag intervals
+Selection process is explained in [Appendix 3](#appendix3), we have explicitly made two model useing the same set of variables so comparison is possible.
+
+### Including only swing states
+As we stated above, the election is determined almost solely by swing states.
+And many other states do not have enough data for sound prediction.
+Thus our model only include the data for swing states and nation polls.
+
+## Model summary
+[@tbl-modelharris] and [@tbl-modeltrump] are the model coefficients fitted for Harris and Trump.
+The model for Harris have adjusted R-squared 0.1781, and model for trump have adjusted R-squared 0.3195.
+
+Note that DJIA_lag_s is the variable $D_{t-s}$
+
+```{r}
+#| label: tbl-modelharris
+#| tbl-cap: Model summary for Harris
+#| include: true
+#| echo: false
+#| warning: false
+#| message: false
+
+model_summary_harris <- summary(model_harris)
+
+coef_table_harris <- as.data.frame(model_summary_harris$coefficients)
+
+kable(coef_table_harris)
+```
+
+```{r}
+#| label: tbl-modeltrump
+#| tbl-cap: Model summary for Trump
+#| include: true
+#| echo: false
+#| warning: false
+#| message: false
+
+model_summary_trump <- summary(model_trump)
+
+coef_table_trump <- as.data.frame(model_summary_trump$coefficients)
+
+kable(coef_table_trump)
+```
+
+# Results {#sec-results}
+
+## Model prediction
+
+[@tbl-predictionharris] shows the predicted result for the election in the swing states as well as the national result. However, this table have not considered confidence intervals.
+Note that the reason for using confidence interval instead of prediction intervals is that as explained in [Use of mean to predict election](@usemean) section, the actual result of the election is expected to be close to the population mean.
+For instance in North Carolina, the 95% confidence interval for Trump is $[48.22990, 49.91708]$, while for Harris it is $[48.90044, 50.54273]$, they have overlapped, which means that we are not very certain with the result.
+To gain a better understanding of the possibility of a person winning in each state and the nation, we use Monte Carlo simulation to get the possibility.
+The last column of [@tbl-predictionharris] is made using this method and give the possibility we predict for Trump to win each state.
+
+```{r}
+#| label: tbl-predictionharris
+#| tbl-cap: Prediction for Each state
+#| include: true
+#| echo: false
+#| warning: false
+#| message: false
+
+graphing_result = prediction_results %>% select(state, prediction_trump,prediction_harris,result,trump_win_rate)
+
+kable(graphing_result, col.names = c("State" , "Prediction Trump" ,"Prediction Harris", "Predicted Result","Trump's Win Rate"))
+```
+
+Note that the United States uses Electoral College system to determine the winner of the election.
+Given this model, we may predict using Monte Carlo simulation the win rate of each candidate.
+Assume that each state except for swing states voted for the the party they voted for in 2020, this gives Democratic party 226 electoral votes and The republican 188 electoral votes.
+Using our model, we find that republican wins 66199 out of 1000000 simulation with no draws, mean electoral votes for democratic is 284.765953 and mean electoral votes for republican is 253.234047.
+The estimated win rate for republican is 0.066199.
+If we follow our prediction with out considering possibility, Harris will win by 293 versus 245 electoral votes.
+
+# Discussion{#sec-discussion}
+Our studies have presented a forecast of the U.S. 2024 presidential election
+
+## Interpretation of Model Coefficients
+First we note that unlike our initial instinct, the votes of the two parties are not always mutually exclusive.
+In the case of $\beta_6$ and $\alpha_6$(DJIA_lag_180), the two candidates both have positively correlated support rate.
+This may indicate that people are less likely to select an independent candidate when the economic is good, as people will be less likely to seek change under such circumstance.
+However in the case of states, we see that every $\beta_s$ have opposite sign from $\alpha_s$.
+
+We also see there is a difference in the R-squares, this may tell us that economy is a better predictor for support rate of Trump than Harris, possibly indicating that Trump has a campaign more focused on economy.
+$\beta_6$ is almost 2 times as large as $\alpha_6$, which aligns with our hypothesis that good long term economy will make the election favor the ruling party.
+
+## Comparson with 2020 
+Looking back at 2020, based on our prediction, the difference was that Harris is predicted to take North Carolina and the entire Maine, While Trump is predicted to win Arizona and Georgia, which makes the republican win 10 more votes than the 2020 election.
+Or 13 more votes after considering the slight changes to number of electoral votes of each state.
+However the Democratic party is in a great advantage and will not be defeated unless Trump wins in North Carolina.
+Which by our prediction is difficult but not impossible.
+
+## Impact of Each Swing state
+The modeling have highlighted the impact swing states has on presidential election.
+And also suggest the definition of swing state is subject to change.
+Florida may be considered a swing state from the election results in 2020, however the polling result and model analysis have shown that Trump will certainly win this state.
+We also need to highlight the importance and closeness of the election in North Carolina.
+As a swing state with 16 electoral votes, we predict Republican Party to have a reasonable probability to win North Carolina, and North Carolina will be deceive in determining the election results.
 
 ## Weaknesses and next steps
+Our studies have relied solely on DJIA data and have not considered other factors that also has great influence to the result of the presidential election.
+Although as Pasek(2015) says, it is never possible to model all the factors, it is even unrealistic to model all the important factors since each year's election is unique in it's way.
+We still may improve the analysis by giving consideration to more variables.
+For instance, we may do a weighted regression by using the numerical scores as weights, or consider other measurements of economy that covers the transportation and utility aspect of economy that DJIA did not cover.
+
+Although we have claimed that non-swing states will almost certainly vote for the party they supported, bringing in their data still may help our.
+However we have to note having all the states will greatly increase the complexity of the model, and prediction for many of the states will not be trustworthy due to scarcity of relevant data.
+
+In the modeling, we have view polling as taking one sample form the population, while election as taking 60% of the data form the population.
+However, polls may be biased by various factors.For example non-response bias or moderacy bias[@stantcheva2023how].
+A closer and more careful investigation by looking at the difference between poll results an actual election outcomes is required to understand the relationship between polls and actual election.
 
@@
 
-# Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-## Diagnostics
-
 \newpage
 
-# Appendix 1: Polling methodology
+# Appendix 1: Polling methodology {#Appendix1}
 
@@
 This will also cause a slight bias in the response.
+
+\newpage
 
@@
 
+# Appendix 3: Model diagnostics {#appendix3}
+
+[@fig-diagnostics_harris] and [@fig-diagnostics_trump] are the model diagnostics for the two models.
+We can see that the residue only have limited deviation from normality and did not display an obvious pattern, which justifies the assumption of linearity and independence.
+
+```{r}
+#| label: fig-diagnostics_harris
+#| tbl-cap: Model diagnostics for Harris model
+#| include: true
+#| echo: false
+#| warning: false
+#| message: false
+par(mfrow = c(2, 2))  
+plot(model_harris)
+```
+
+```{r}
+#| label: fig-diagnostics_trump
+#| tbl-cap: Model diagnostics for Trump model
+#| include: true
+#| echo: false
+#| warning: false
+#| message: false
+par(mfrow = c(2, 2))  
+plot(model_trump)
+```
+## Selection
+
+Our model selection is based on AIC(Akaike information criterion) and interpretation of data.
+
+The best model in terms of AIC in selecting was "DJIA_lag_7 + DJIA_lag_14 + DJIA_lag_90 + DJIA_lag_120 + DJIA_lag_150 + DJIA_lag_180 + state" for Harris and "DJIA_lag_14 + DJIA_lag_28 + DJIA_lag_60 + DJIA_lag_90 + DJIA_lag_150 + DJIA_lag_180 + state" For Trump.
+However, we want the two model to have the same set of predictors, which may increase interpretability of the model.
+So we have selected a model that include relevant terms for both candidates.
+
+\newpage
+
 # References

### 2024-11-04 scripts/06-model_predict.R
Added script for prediction

@@ -1,0 +1,126 @@
+#### Preamble ####
+# Purpose: Test various models and choose one to save
+# Author: David Qi
+# Date: Nov Oct 2024
+# Contact: david.qi@mail.utoronto.ca 
+# License: MIT
+# Pre-requisites: Model are complete and saved in "models/", cleaned DJIA data saved in "data/02-analysis_data/DJIA_filled.parquet"
+# Any other information needed? No
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(arrow)
+
+model_harris = readRDS(file = here::here("models/model_harris.rds"))
+
+model_trump = readRDS(file = here::here("models/model_trump.rds"))
+
+DJIA_data = read_parquet("data/02-analysis_data/DJIA_filled.parquet")
+
+# Set the target date
+target_date <- as.Date("2024-11-05")
+
+
+# States to predict
+states_to_predict <- c("North Carolina", "Pennsylvania", "Georgia", 
+                       "Arizona","Nevada", "Michigan", "Wisconsin", "Maine CD-2", "Florida", 
+                       "National")
+
+
+
+DJIA_values <- DJIA_data %>% filter(DATE %in% c(
+  target_date - 7,
+  target_date - 14,
+  target_date - 28,
+  target_date - 60,
+  target_date - 150,
+  target_date - 180
+))
+
+
+  # Create a new data frame for the current state
+new_data <- data.frame(
+    DJIA_lag_7   = rep(DJIA_values$DJIA[6],10),
+    DJIA_lag_14  = rep(DJIA_values$DJIA[5],10),
+    DJIA_lag_28  = rep(DJIA_values$DJIA[4],10),
+    DJIA_lag_60  = rep(DJIA_values$DJIA[3],10),
+    DJIA_lag_150 = rep(DJIA_values$DJIA[2],10),
+    DJIA_lag_180 = rep(DJIA_values$DJIA[1],10),
+    state        = states_to_predict
+  )
+  
+
+predictions_trump <- predict(model_trump, newdata = new_data,interval="confidence")
+
+predictions_harris <- predict(model_harris, newdata = new_data,interval="confidence")
+
+
+#### Monte Carlo simulation ####
+set.seed(304)
+
+predictions_trump_se <- predict(model_trump, newdata = new_data,se.fit = TRUE)
+predictions_harris_se <- predict(model_harris, newdata = new_data,se.fit = TRUE)
+standard_errors_trump <- predictions_trump_se$se.fit
[…]

### 2024-11-04 paper/paper.qmd
Change wording, correct typo

@@
 The study may offer guidance for campaign strategies and market forcasts.
-While providing insight on changes of swing states in U.S.
-politics.
-"
+While providing understanding on changes of swing states in U.S.
+politics and how Florida is no longer a swing state."
 format: pdf
@@
 Along with the votes that we predict he will almost certainly gain, this pushes him one step (9 electoral votes) to victory.
-This provide power insight to how campaign and fund raising strategies should be targeted on the election day.
+This provide powerful guidence to how campaign and fund raising strategies should be targeted on the election day.
 And provides a forecast on one of the most important influences in economy in the next several months, which will be especially helpful in planning market stragities.
@@
 An alternative approach is to use changes in the DJIA to reflect economic fluctuations.
-However, we choose to use the absolute DJIA values, as inputing missing days for change data is difficult to justify, and the absolute economic status provides essential insights.
+However, we choose to use the absolute DJIA values, as inputing missing days for change data is difficult to justify, and the absolute economic status cannot be reflected by the change in one day.
 Our hypothesis is that a strong economy favors the ruling party in elections, as switching parties could risk economic shifts and potentially lower living standards.

### 2024-11-04 scripts/06-model_predict.R

@@ -1,7 +1,7 @@
 #### Preamble ####
-# Purpose: Test various models and choose one to save
+# Purpose: make prediction using the model
 # Author: David Qi
-# Date: Nov Oct 2024
+# Date: 04 Nov 2024
 # Contact: david.qi@mail.utoronto.ca 
 # License: MIT
 # Pre-requisites: Model are complete and saved in "models/", cleaned DJIA data saved in "data/02-analysis_data/DJIA_filled.parquet"

### 2024-11-04 paper/paper.qmd
Correct graph and typo in appendix 3

@@
 
-##Data Measurement
+## Data Measurement
 The data was organized by FiveThirtyEight (2024), collected form polls made by various institutions, they each use a distinct methodology.
@@
 #| label: fig-samplesize
-#| tbl-cap: Comparison of average polling result for two candiates
+#| fig-cap: Sample Sizes of the Polls We Use
 #| include: true
@@
 
-# Appendix 1: Polling methodology {#Appendix1}
+# Appendix 1: Polling methodology {#appendix1}
 
@@
 ## Sampling Approach
-This survey employs Stratified Random Sampling, which involves divide the target population(In this case, the people who are eligible to vote in US) into subgroups(strata) based on their demographics characteristic[@stantchevaonline].
+This survey employs Stratified Random Sampling, which involves divide the target population(In this case, the people who are eligible to vote in US) into subgroups based on their demographics characteristic [@stantchevaonline].
 To simulate the national electorate, stratify the sample according to age, gender, race/ethnicity, education level, urban/rural domicile, and region.
@@
 ## Respondent Recruitment
-Partner with an online panel provider to find respondents that match each stratum’s specifications[@stantchevaonline].
+Partner with an online panel provider to find respondents that match each stratum’s specifications [@stantchevaonline].
 Specify quotas to ensure that, for instance, you wish to have 18% of respondents from the Midwest or 25% aging between 30 - 40.
@@
 ## Data Validation
-To keep distinct replies between responses, make sure respondents don't participate more than once each wave, and highlight straight-lining, inconsistent replies, and other low-effort answers for data quality purposes[@horn1997effective].
+To keep distinct replies between responses, make sure respondents don't participate more than once each wave, and highlight straight-lining, inconsistent replies, and other low-effort answers for data quality purposes [@horn1997effective].
 
@@
 #| label: fig-diagnostics_harris
-#| tbl-cap: Model diagnostics for Harris model
+#| fig-cap: Model diagnostics for Harris model
 #| include: true
@@
 #| message: false
+#| fig.width: 10
+#| fig.height: 7
+
 par(mfrow = c(2, 2))  
@@
 #| label: fig-diagnostics_trump
-#| tbl-cap: Model diagnostics for Trump model
+#| fig-cap: Model diagnostics for Trump model
 #| include: true

### 2024-11-04 paper/paper.qmd

@@
 From our prediction, we highlight North Carolina as the most important battleground state.
-The state has a predicted $14.2\% $probability to be won by Trump, and will give him 16 electoral votes.
+The state has a predicted $14.2\%$ probability to be won by Trump, and will give him 16 electoral votes.
 Along with the votes that we predict he will almost certainly gain, this pushes him one step (9 electoral votes) to victory.
@@
 
-The scripts in this paper are written in R[@citeR], The table in our study are generated by knitr [@knitr],  and graph with ggplot2[@ggplot2].
+The scripts in this paper are written in R[@citeR],  using mainly tidyverse[@tidyverse] The table in our study are generated by knitr [@knitr],  and graph with ggplot2[@ggplot2].
 Data input and output were done using arrow[@arrow], and curl[@curl].
